DOM bridges and barriers: systematic review and survey

### 2025-09-22 reports/draft_results.qmd
updated figures

**Which themes overlap between the literature and surveys?**

**Which themes only appear in the literature? Which themes are unique to the surveys?**

@@

::: panel-tabset

#### Review: Bridge themes

@@

::: panel-tabset

### Literature 
Code appearance frequencies across articles (n = 118). The proportions reported in these figures represent the percent of papers that mentioned a subcode within these broader code categories at least once.

@@

#plot
br_freq <- ggplot(code_sum_br, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "goldenrod") + 
  scale_x_continuous(limits = c(0, 30), 
                     breaks = seq(0, 30, 5), 
                     name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

@@

#plot
ba_freq <- ggplot(code_sum_ba, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  scale_x_continuous(limits = c(0, 30), 
                     breaks = seq(0, 30, 5), 
                     name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

@@

![](images/br_ba_freq.png){fig-align="center" width="8in", height = "10in"}

@@

:::

## Result 3: Variations by industry (fishing vs. shipping)

>>>>

**Which themes appear in both sources?**

**What new information have surveys added beyond what was already written in the literature?**

@@

::: panel-tabset
#### Review: Bridge themes

@@

Code appearance frequencies across all bridge **OR** barrier codes. The proportions reported in these figures represent the percent of codes that theme or sub-theme comprises.

::: panel-tabset
### Literature

@@

#plot
br_freq <- ggplot(code_sum_br, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "goldenrod") + 
  # scale_x_continuous(limits = c(0, 30), 
  #                    breaks = seq(0, 30, 5), 
  #                    name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  scale_x_reverse(limits = c(30, 0), 
                  #breaks = seq(0, 30, 5), 
                  name = "Percent appearance (%)")+
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

@@

#plot
ba_freq <- ggplot(code_sum_ba, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  scale_x_continuous(limits = c(0, 30), 
                     breaks = seq(0, 30, 10), 
                     name = "Percent appearance (%)") +
  scale_y_discrete(position = "right") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 1), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

@@

![](images/br_ba_freq.png){fig-align="center" width="10in", height = "12in"}

@@

#theme bridge/barrier
code_sum_s <- code_freq_s %>%
  group_by(bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(code_cat) %>% 
  mutate(n_subtotal = sum(total)) %>%
  summarise(value = n_subtotal/n_total) %>%
  rename("br_ba" = code_cat) %>%
  mutate(percent_label = scales::percent(value, accuracy = 1), 
         percent = if_else(br_ba == "Data barrier" | br_ba == "Ecological barrier" | br_ba == "Social barrier" | br_ba == "Implementation barrier", value, -value),
         code_level = if_else(br_ba %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
                              if_else(br_ba %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
                              if_else(br_ba %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", 
                              if_else(br_ba %in% c("Implementation bridge", "Implementation barrier"), "Implementation bridge or barrier", NA)))), 
         code_brba = if_else(br_ba %in% c("Data bridge", "Ecological bridge", "Social bridge", "Implementation bridge"), "bridge", "barrier")) %>%
  group_by(code_level) %>%
  distinct()

code_sum_s$code_level <- factor(code_sum_s$code_level, levels = c("Implementation bridge or barrier", "Social bridge or barrier", "Ecological bridge or barrier", "Data bridge or barrier"))
code_sum_s$code_brba <- factor(code_sum_s$code_brba, levels = c("bridge", "barrier"))

all_freq_survey <- ggplot(code_sum_s, aes(x = code_level, 
             y = percent,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = percent_label),
            position = position_stack(vjust = 0.65),
            color = "black", 
            size = 5) +
  ylim(-0.55,0.7)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"), labels = c("Bridge frequency", "Barrier frequency")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 26, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.17, 0.99),
        legend.text = element_text(size = 20, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 18)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))

#ggsave("figures/ms_figs/all_freq/all_freq_survey.png", width = 13, height = 8, units = ("in"))

#bridge freq
#reformat data
code_freq_s$code_cat <- factor(code_freq_s$code_cat, levels = c("Data bridge", "Data barrier", "Ecological bridge", "Ecological barrier", "Social bridge", "Social barrier", "Implementation bridge","Implementation barrier"))

code_sum_br_s <- code_freq_s %>%
  filter(bridge_barr == "Bridge") %>%
  mutate(n_total = sum(total)) %>%
  group_by(code_cat, code_name) %>%
  summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(percent))

#plot
br_freq_s <- ggplot(code_sum_br_s, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "goldenrod") + 
  # scale_x_continuous(limits = c(0, 30), 
  #                    breaks = seq(0, 30, 5), 
  #                    name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  scale_x_reverse(limits = c(20, 0), 
                  #breaks = seq(0, 30, 5), 
                  name = "Percent appearance (%)")+
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/br_freq_survey.png", width = 6, height = 8, units = c("in"))

#barrier freq
#reformat data
code_sum_ba_s <- code_freq_s %>%
  filter(bridge_barr == "Barrier") %>%
  mutate(n_total = sum(total)) %>%
  group_by(code_cat, code_name) %>%
  summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(percent))

#plot
ba_freq_s <- ggplot(code_sum_ba_s, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  scale_x_continuous(limits = c(0, 20), 
                     breaks = seq(0, 20, 10), 
                     name = "Percent appearance (%)") +
  scale_y_discrete(position = "right") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 16, hjust = 1), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/ba_freq_survey.png", width = 6, height = 9, units = c("in"))

![](images/br_ba_freq_survey.png){fig-align="center" width="10in", height = "12in"}
:::

## Result 3: Variations by industry

### 3.1: Conservation problems DOM has been used to address

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-height: 5.5
#| fig-align: center

codes_vars$broad_location <- as.factor(codes_vars$broad_location)
codes_LU <- codes_vars[!duplicated(codes_vars$Lit_unit), ]

consv <- codes_LU %>%
  gather(col, consv_topic, reason1:reason3) %>%
  subset(select = consv_topic)

consv_s <- consv %>%
  group_by(consv_topic) %>%
  filter(consv_topic != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

consv_bar <- ggplot(consv_s) +
  geom_bar(aes(x = perc, y = reorder(consv_topic, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 45))+
  scale_y_discrete(expand = expansion(add = c(0,0))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Arial", size = 12)) + 
  geom_text(aes(x = perc, y = consv_topic, label = consv_topic),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Arial",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Arial",
      size = 13)
  )

consv_bar
```

### 3.2: Fishing vs. shipping
